sem

### 2023-03-20 scripts/3_sem_visualization.qmd
first revisions to ncc 2023-03-17

```{r}
library(lavaan)

@@

@@

```{r}
samples <- here::here("data", "samples", "bec") %>%
  list.files(full.names = T, pattern = ".shp$") %>%
  map(read_sf) %>%
  bind_rows()

samples %>%
  st_drop_geometry() %>%
  count(zone) %>%
  ggplot(aes(
    x = n,
    y = fct_reorder(zone, n),
    fill = zone
  )) +
  geom_col() +
  ggsci::scale_fill_igv() +
  theme_bw() +
  labs(x = "n",
       y = NULL,
       fill = NULL)
```

```{r}
all_rasts_loc <- here::here("data", "rasts", "all_rasts.dat")

if (!file.exists(all_rasts_loc)) {
  struct_locs <- here::here("G://", "mosaiced", "structure")
  struct_rasts <- list.files(
    struct_locs,
    recursive = T,
    pattern = ".dat$",
    full.names = T
  ) %>%
    map(rast) %>%
    rast()

@@

  dhi_rasts <-
    list.files(here::here("G://", "mosaiced", "DHI"),
               pattern = "2010s.tif$",
               full.names = T) %>%
    map(rast) %>%
    rast()

@@

  dhi_mm <- minmax(dhi_rasts)

  dhi_rasts_mm <- (dhi_rasts - dhi_mm[1,]) / (dhi_mm[2,] - dhi_mm[1,])

  all_rasts <- c(vlce, struct_rasts, dhi_rasts_mm)

  writeRaster(all_rasts, all_rasts_loc, overwrite = T, filetype = "envi",
              datatype = "FLT4S")
}

all_rasts <- rast(all_rasts_loc)
```

```{r}
extracted_samples <- extract(all_rasts, samples, bind = T)
```

>>>>

```{r}
library(lavaan)
library(lavaanPlot)
library(semPlot)

@@

# normalization from grace et al 2016
# first does mylog (does natural logarithm. if 0s present, does log(one + x))
# then scales to z scores
# results it more normal looking data
normal <- function(data) {
  data %>%
    mylog() %>%
    scale() %>%
    as.numeric()
}
```

@@

```{r}
bec_samples <- here::here("data", "samples", "bec") %>%
  list.files(full.names = T, pattern = ".shp$") %>%
  map(read_sf) %>%
  bind_rows()

vlce_samples <- here::here("data", "samples", "vlce") %>%
  list.files(full.names = T, pattern = ".shp$") %>%
  map(read_sf) %>%
  bind_rows()
```

```{r}

all_rasts_loc <- here::here("data", "rasts", "all_rasts_bec.dat")

if (!file.exists(all_rasts_loc)) {
  # bec_rast <- bec() %>%
  #   rasterize(y = vlce, field = "ZONE") %>%
  #   crop(bcb_rast) %>%
  #   mask(bcb_rast)
  # 
  # names(bec_rast) <- "bec"

  struct_locs <- here::here("F://", "mosaiced", "structure")
  struct_rasts <- list.files(
    struct_locs,
    recursive = T,
    pattern = ".dat$",
    full.names = T
  ) %>%
    map(rast) %>%
    rast()

@@

  dhi_rasts <-
    list.files(here::here("F://", "mosaiced", "DHI"),
               pattern = "2010s.tif$",
               full.names = T) %>%
    map(rast) %>%
    rast()

@@

  # dhi_mm <- minmax(dhi_rasts)
  #
  # dhi_rasts_mm <- (dhi_rasts - dhi_mm[1,]) / (dhi_mm[2,] - dhi_mm[1,])

  all_rasts <- c(vlce, struct_rasts, dhi_rasts) %>%
    mask(bcb_rast)

  writeRaster(
    all_rasts,
    all_rasts_loc,
    overwrite = T,
    filetype = "envi",
    datatype = "FLT4S"
  )
}

all_rasts <- rast(all_rasts_loc)

bec_rast <- bec() %>%
  rasterize(y = vlce, field = "ZONE") %>%
  crop(bcb_rast) %>%
  mask(bcb_rast)

names(bec_rast) <- "bec"
all_rasts <- c(bec_rast, vlce, all_rasts)
```

```{r}
bec_sample_loc <- here::here("data", "samples", "bec_extracted.shp")
vlce_sample_loc <- here::here("data", "samples", "vlce_extracted.shp")

if (!file.exists(bec_sample_loc)) {
  bec_samples_e <- extract(all_rasts, bec_samples, bind = T)
  writeVector(bec_samples_e, bec_sample_loc, overwrite = T)
}
if (!file.exists(vlce_sample_loc)) {
  vlce_samples_e <- extract(all_rasts, vlce_samples, bind = T)
  writeVector(vlce_samples_e, vlce_sample_loc, overwrite = T)
}

bec_samples_e <- vect(bec_sample_loc)
vlce_samples_e <- vect(vlce_sample_loc)
```

bec_df <- bec_samples_e %>% 
  st_as_sf() %>%
  st_drop_geometry() %>%
  select(bec:VarDHI) %>%
  drop_na() %>%
  tibble() %>%
  filter(!(bec %in% invalid_zones))

names_manual <- c(
  "bec",
  "category",
  "basal_area",
  "elev_cv",
  "elev_mean",
  "elev_p95",
  "elev_stddev",
  "gross_stem_volume",
  "loreys_height",
  "percentage_first_returns_above_2m",
  "percentage_first_returns_above_mean",
  "total_biomass",
  "CumDHI",
  "MinDHI",
  "VarDHI"
)

# names(bec_df) <- names(all_rasts)
names(bec_df) <- names_manual

```

```{r}

bec_df <- bec_df %>%
  mutate(strata = bec) %>%
  relocate(strata)

vlce_df <- vlce_samples %>%
  st_as_sf() %>%
  st_drop_geometry() %>%
  select(bec:VarDHI)

# names(bec_df) <- names(all_rasts)
names(vlce_df) <- names_manual

vlce_df <- vlce_df %>%
  mutate(strata = category) %>%
  relocate(strata)

strata_df <- bind_rows(bec_df, vlce_df) %>%
  filter(!(loreys_height < elev_mean | elev_p95 < elev_mean)) %>%
  pivot_longer(basal_area:total_biomass, names_to = "variable") %>%
  left_join(keys$continuous) %>%
  mutate(value = value / divide_by) %>%
  select(strata, bec, category, ends_with("DHI"), variable, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  unchop(everything()) %>% # somehow does the fix i want
  filter(elev_cv <= 1) %>% # remove erroneous values
  drop_na() # remove anything w/ NA

# groups by zone and all other structure variables 
# takes the mean of each DHI dataset
# effectively removing duplicates, but takes the mean instead of the first value
strata_df_b <- strata_df %>%
  group_by(across(c(strata, bec, category, basal_area:total_biomass))) %>%
  dplyr::summarize(across(ends_with("DHI"), mean)) %>%
  ungroup() %>%
  filter(!(bec %in% invalid_zones))
```

```{r}

# see number of pixels
n_dups <- strata_df %>%
  group_by(across(c(
    strata, bec, category, basal_area:total_biomass
  ))) %>%
  summarize(n = n())

egg <- n_dups %>%
  filter(!(bec %in% invalid_zones)) %>%
  group_by(strata, duplicated = n != 1) %>%
  summarize(duplicate_pixels = sum(n)) %>%
  pivot_wider(names_from = duplicated, 
              values_from = duplicate_pixels) %>%
  mutate(total = `TRUE` + `FALSE`,
         per_dup = `TRUE` / total,
         strata_name = glue::glue("{strata} (n = {total})"),
         label = glue::glue(" n[dup]: {total}"))

n_dups %>% 
  ggplot(aes(x = n, fill = strata)) + 
  geom_histogram() + 
  facet_wrap(~strata, scales = "free")

egg %>%
  ggplot(aes(x = fct_reorder(strata_name, per_dup), 
             y = per_dup)) +

  geom_col() +
  geom_text(aes(label = label), hjust = 0,
            parse = T) + 
  lims(y = c(0, 1)) +
  labs(x = NULL,
         y = "Proportion of Duplicates") +
  coord_flip() +
  theme(panel.grid = element_blank())

n_dups %>%
  ggplot(aes(x = n)) +
  geom_histogram() +
  labs(x = "n",
       y = NULL,
       title = "Number of duplicated structure pixels") +
  theme_light()

```

```{r}
# downsampling procedure - allows for comparisons between models if 
# sample size is the same (AIC and BIC are based on sample)

downsample_no <- strata_df_b %>%
  count(strata) %>%
  pull(n) %>%
  min()
```

```{r}
set.seed(69420)
# downsample the vlce with equivalent sampling across becs
# if possible, if not it takes the maximum available samples and adjusts to 
# account

downsample_prop <- function(strat_tib) {
  # this function samples equal to the downsample number
  # across each bec zone, for each strata
  # if a zone does not have enough to be an equal number of samples
  # it takes all of those samples, and adjusts other zones to meet
  # the total number of desired samples (downsample number)
  bec_counts <- strat_tib %>%
    count(bec)

  total_bec <- bec_counts %>%
    pull(bec) %>%
    length()

  bec_sub <- bec_counts %>%
    filter(n <= ceiling(downsample_no / 12)) %>%
    pull(bec)

  df_sub <- strat_tib %>%
    filter(bec %in% bec_sub)

  n_sub <- nrow(df_sub)

  n_rest <- downsample_no - n_sub

  df_over <- strat_tib %>%
    filter(!(bec %in% bec_sub)) %>%
    group_by(bec) %>%
    slice_sample(n = ceiling(n_rest / (total_bec - length(bec_sub)))) %>%
    ungroup() %>%
    slice_sample(n = n_rest)

  bind_rows(df_over, df_sub)
}

vlce_ds <- strata_df_b %>%
  filter(strata == category) %>%
  group_by(strata) %>%
  group_split() %>%
  map(downsample_prop) %>%
  bind_rows()

# downsample bec zones. does not matter which forest type they come from

bec_ds <- strata_df_b %>%
  filter(strata != category) %>%
  group_by(strata) %>%
  slice_sample(n = downsample_no)

strata_ds <- bind_rows(vlce_ds, bec_ds)

meg_dir <- here::here("data", "samples", "meg")

dir.create(meg_dir, showWarnings = F)
write_csv(strata_ds, here::here(meg_dir, "downsampled.csv"))
write_csv(strata_df_b, here::here(meg_dir, "full.csv"))
```

```{r}
clean_df <- strata_ds %>%
  mutate(
    percentage_first_returns_above_mean = percentage_first_returns_above_mean,
    understory_per = percentage_first_returns_above_2m - percentage_first_returns_above_mean,
    percentage_first_returns_above_2m = percentage_first_returns_above_2m,
    overstory_range = loreys_height - elev_mean,
    understory_range = elev_mean - 2,
    over_biomass = total_biomass * (percentage_first_returns_above_mean / percentage_first_returns_above_2m),
    under_biomass = total_biomass * (understory_per / percentage_first_returns_above_2m)
  ) %>%
  relocate(strata,
           bec,
           category,
           ends_with("DHI"))

# applies normalization by strata of analysis (i.e. bec zone or forest type)

normal_df <- clean_df %>%
  group_by(strata) %>%
  mutate(across(CumDHI:under_biomass, .fns = ~ normal(.x)))

write_csv(normal_df, here::here(meg_dir, "downsampled_normal.csv"))
```

```{r}
dhi_vars <- c("CumDHI",
              "VarDHI",
              "MinDHI")
models <- c(
  paste0(
    dhi_vars,
    " ~ elev_p95 + total_biomass + elev_cv
    elev_cv ~ elev_p95
    total_biomass ~ elev_p95"
  ),
  paste0(
    dhi_vars,
    " ~ percentage_first_returns_above_2m + basal_area + percentage_first_returns_above_mean
    percentage_first_returns_above_mean ~ percentage_first_returns_above_2m
    basal_area ~ percentage_first_returns_above_2m"
  )
)

splits <- normal_df %>%
  group_by(strata) %>%
  group_split()
```

```{r}
lavaans <- crossing(data = splits, models = models) %>%
  mutate(outs = map2(.x = models, .y = data, .f = sem))

fitmeasures_list <- c("chisq", "df", "pvalue", "cfi", "rmsea", "aic", "srmr")

myfits <- function(lavaan) {
  try(fitmeasures(lavaan, fit.measures = fitmeasures_list), silent = T)
}

tib_out <- lavaans %>%
  mutate(var = word(models, start = 1),
         strata = data %>% map(pull, var = strata) %>% map(unique) %>% unlist(),
         fits = outs %>% map(myfits),
         pars = outs %>% map(parameterestimates) %>% map(tibble)#,
         # interpreted = outs %>% map(effectsize::interpret)
         ) %>%
  relocate(data, models)

non_converge <- tib_out %>%
  filter(fits %>% map(class) == "try-error")

converge <- tib_out %>%
  # remove any models that don't converge
  filter(!(fits %>% map(class) == "try-error"))

converge_fm <- converge %>%
  select(models, var, strata, fits, pars) %>%
  mutate(fits = fits %>% map(as.numeric) %>%
           map(tibble) %>%
           map(mutate, measures = fitmeasures_list) %>%
           map(rename, value = `<dbl>`) %>%
           map(pivot_wider, names_from = measures, values_from = value)) %>%
  unnest(fits) %>%
  mutate(mod_follows = str_remove(models, ".*DHI"))
```

```{r}
global_fits <- converge_fm %>%
  mutate(global_sig = pvalue > 0.05,
         mod_follows = str_remove(models, ".*DHI")) %>%
  group_by(mod_follows) %>%
  dplyr::summarize(per_sig = (sum(global_sig) / n()),
                   count = n()) %>%
  arrange(desc(per_sig))

global_fits

# fan et al. (2016) says chi ^ 2 saturates w/ samples size and as such isn't
# the be all end all
# need to also report: SRMR, CFI, and RMSEA
# aic can be used to compare models
# CFI > 0.95
# RMSEA < 0.06
# SRMR < 0.09
```

```{r}
vlce_colours <- keys$vlce %>%
  filter(forest == "Forest") %>%
  select(strata = name_clean, colour = ntems_colour)

# bc government bec colours.
# i prefer the igv palette
bec_colours <- bcmaps::bec_colors() %>%
  tibble(strata = names(bcmaps::bec_colours()), colour = .)

bec_colours <- keys$bec %>%
  mutate(colour = ggsci::pal_igv("default")(16)) %>%
  select(strata = zone, colour) %>% 
  filter(!(strata %in% invalid_zones))

strata_colours <- bind_rows(vlce_colours, bec_colours) %>%
  arrange(strata)

mod_names <- c("Vertical Structure",
               "Horizontal Structure")

selected_models <- models %>%
  str_remove(".*DHI") %>%
  unique()

names_df <- tibble(mod_follows = selected_models,
                   mod_names,
                   order = 1:2)

converge_fm %>%
  mutate(mod_follows = str_remove(models, ".*DHI")) %>%
  left_join(names_df) %>%
  ggplot(aes(x = pvalue, y = chisq, col = strata)) +
  geom_point() +
  geom_vline(xintercept = 0.05,
             col = "red",
             lty = "dashed") +
  annotate(
    "Text",
    label = "Global Fit Accepted",
    x = 0.051,
    y = 100,
    angle = 270,
    col = "red"
  ) +
  facet_wrap(~ fct_reorder(mod_names, order),
             ncol = 1) +
  scale_colour_manual(values = strata_colours %>% pull(colour) %>% unname()) +
  theme_bw() +
  labs(x = "P-value",
       y = (expression(chi^2)),
       colour = NULL) +
  theme(panel.grid = element_blank())
```

```{r}
par_df <- converge_fm %>%
  group_by(strata) %>%
  select(models, var, strata, pars, mod_follows) %>%
  unnest(cols = pars) %>%
  filter(op == "~")

split_dfs <- par_df %>%
  group_by(strata) %>%
  group_split()

fit_df <- converge_fm %>%
  select(models, var, strata, mod_follows, pvalue:srmr)

coords_df <-
  tibble(
    pos = c(
      "elev_p95",
      "elev_cv",
      "total_biomass",
      "percentage_first_returns_above_2m",
      "basal_area",
      "percentage_first_returns_above_mean",
      "dhi",
      "dhi"
    ),
    x = c(0, 2, 2, 0, 2, 2, 4, 4),
    y = c(0, 1,-1, 0, 1,-1, 0, 0),
    mod_follows = c(
      selected_models[1],
      selected_models[1],
      selected_models[1],
      selected_models[2],
      selected_models[2],
      selected_models[2],
      selected_models[1],
      selected_models[2]
    )
  )

arrows_df <-
  tibble(
    rhs = c(
      "elev_p95",
      "elev_cv",
      "total_biomass",
      "elev_p95",
      "elev_p95"
    ),
    lhs = c(
      "dhi",
      "dhi",
      "dhi" ,
      "elev_cv",
      "total_biomass"
    )
  ) %>%
  left_join(coords_df, by = c(lhs = "pos")) %>%
  rename(xend = x,
         yend = y) %>%
  left_join(coords_df, by = c(rhs = "pos")) %>%
  rename(xstart = x,
         ystart = y) %>%
  mutate(
    xend = xend * 0.975,
    yend = yend * 0.9,
    yend = case_when(ystart == 1 ~ 0.15,
                     ystart == -1 ~ -0.15,
                     T ~ yend),
    ystart = ystart * 0.9
  )

arrows_df2 <-
  tibble(
    rhs = c(
      "percentage_first_returns_above_2m",
      "basal_area",
      "percentage_first_returns_above_mean",
      "percentage_first_returns_above_2m",
      "percentage_first_returns_above_2m"
    ),
    lhs = c(
      "dhi",
      "dhi",
      "dhi",
      "basal_area",
      "percentage_first_returns_above_mean"
    )
  ) %>%
  left_join(coords_df, by = c(lhs = "pos")) %>%
  rename(xend = x,
         yend = y) %>%
  left_join(coords_df, by = c(rhs = "pos")) %>%
  rename(xstart = x,
         ystart = y) %>%
  mutate(
    xend = xend * 0.975,
    yend = yend * 0.9,
    yend = case_when(ystart == 1 ~ 0.15,
                     ystart == -1 ~ -0.15,
                     T ~ yend),
    ystart = ystart * 0.9
  )

arrows_both <- bind_rows(arrows_df, arrows_df2) %>%
  rename(mod_follows = mod_follows.x) %>%
  select(-mod_follows.y)

label_df <- coords_df %>% 
  crossing(var = c("CumDHI", "VarDHI", "MinDHI")) %>% 
  mutate(pos = case_when(pos == "dhi" ~ var, T ~ pos)) %>%
  mutate(x = case_when(pos == "elev_p95" | pos == "percentage_first_returns_above_2m" ~ 0.25, T ~ x)) %>%
  left_join(keys$continuous, by = c("pos" = "variable")) %>%
  mutate(var_long = case_when(is.na(var_long) ~ pos,
                              T ~ var_long),
         x = case_when(endsWith(pos, "DHI") ~ 3.85,
                       T ~ x)) %>%
  select(!c(divide_by, unit))

plot_sem <- function(split_df) {
  strata_name <- split_df %>%
    pull(strata) %>%
    unique()

  fit <- fit_df %>%
    filter(strata == strata_name)

  plot_df <- split_df %>%
    mutate(lhs = case_when(endsWith(lhs, "DHI") ~ "dhi", T ~ lhs)) %>%
    left_join(arrows_both) %>%
    mutate(lhs = case_when(lhs == "dhi" ~ var,
                           T ~ lhs))

  est_label_df <- plot_df %>%
    group_by(var, mod_follows) %>%
    mutate(xmid = ((xend + xstart) / 2),
           ymid = ((yend + ystart) / 2) + 0.125) %>%
    select(var, xmid, ymid, est)

  plot_df %>%
    ggplot() +
    geom_segment(
      aes(
        x = xstart,
        y = ystart,
        xend = xend,
        yend = yend,
        lty = pvalue < 0.05,
        col = est,
      ),
      arrow = arrow(),
      linewidth = 2
    ) +
    # geom_segment(aes(x = xstart, y = ystart, xend = xend, yend = yend,
    #                  col = pvalue < 0.05),
    #              ) +
    geom_label(data = label_df, aes(x = x, y = y, label = var_long)) +
    geom_text(data = est_label_df, aes(
      x = xmid,
      y = ymid,
      label = round(est, digits = 2)
    )) +
    facet_grid(rows = vars(var),
               cols = vars(mod_follows)) +
    theme_void() +
    theme(strip.text = element_blank()) +
    labs(lty = "Significant Path",
         col = "Standardized Estimate",
         title = strata_name) +
    theme(legend.position = "bottom") +
    scico::scale_colour_scico(palette = "hawaii",
                              midpoint = 0,
                              limits = c(-2, 2)) +
    scale_linetype_manual(values = c("dashed", "solid")) +
    geom_text(data = fit, aes(
      x = 3.7,
      y = 0.75,
      label = paste0(
        "CFI:",
        round(cfi, digits = 3),
        "\nRMSEA:",
        round(rmsea, digits = 3),
        "\nSRMR:",
        round(srmr, digits = 3),
        "\nP-Value:",
        round(pvalue, digits = 3)
      )
    ))

  ggsave(here::here("outputs", "strata_sem", paste0(strata_name, ".png")), dpi = 300,
         width = unit(13, "inches"),
         height = unit(11, "inches"),
         bg = "white")

}

map(split_dfs, plot_sem)
```

```{r}
# this cell is about aggregating to the thing. i also want to look at it by BEC
# vs forest type

vlce_classes <- vlce_colours %>%
  pull(strata)

sig_est_df <- par_df %>% 
  ungroup() %>%
  mutate(sig_est = ifelse(pvalue < 0.05, est, NA))

all_est_df <- sig_est_df %>%
  group_by(var, lhs, rhs, mod_follows) %>%
  summarize(num_sig = sum(pvalue < 0.05),
            per_sig = num_sig / n(),
            mean_sig_est = mean(sig_est, na.rm = T)) %>%
  arrange(mean_sig_est) %>%
  mutate(strata = "All Strata")

vlce_est_df <- sig_est_df %>%
  filter(strata %in% vlce_classes) %>%
  group_by(var, lhs, rhs, mod_follows) %>%
  summarize(num_sig = sum(pvalue < 0.05),
            per_sig = num_sig / n(),
            mean_sig_est = mean(sig_est, na.rm = T)) %>%
  arrange(mean_sig_est) %>%
  mutate(strata = "Forest Type")

bec_est_df <- sig_est_df %>%
  filter(!(strata %in% vlce_classes)) %>%
  group_by(var, lhs, rhs, mod_follows) %>%
  summarize(num_sig = sum(pvalue < 0.05),
            per_sig = num_sig / n(),
            mean_sig_est = mean(sig_est, na.rm = T)) %>%
  arrange(mean_sig_est) %>%
  mutate(strata = "BEC Zones")
```

```{r}
group_dfs <- list(all_est_df, vlce_est_df, bec_est_df)

plot_sem_group <- function(group_df) {
  strata <- group_df %>%
    pull(strata) %>%
    unique

  plot_df <- group_df %>%
    mutate(lhs = case_when(endsWith(lhs, "DHI") ~ "dhi", T ~ lhs)) %>%
    left_join(arrows_both) %>%
    mutate(lhs = case_when(lhs == "dhi" ~ var,
                           T ~ lhs)) %>%
    ungroup()

  est_label_df <- plot_df %>%
    group_by(var, mod_follows) %>%
    mutate(xmid = (xend + xstart) / 2,
           ymid = ((yend + ystart) / 2) + 0.125) %>%
    select(var, xmid, ymid, mean_sig_est)

  plot_df %>%
    ggplot() +
    geom_segment(aes(
      x = xstart,
      y = ystart,
      xend = xend,
      yend = yend,
      linewidth = per_sig,
      col = mean_sig_est,
    ),
    arrow = arrow()) +
    # geom_segment(aes(x = xstart, y = ystart, xend = xend, yend = yend,
    #                  col = pvalue < 0.05),
    #              ) +
    geom_label(data = label_df, aes(x = x, y = y, label = var_long)) +
    geom_text(data = est_label_df, aes(x = xmid, y = ymid, label = round(mean_sig_est, digits = 2))) +
    facet_grid(rows = vars(var),
               cols = vars(mod_follows)) +
    theme_void() +
    theme(strip.text = element_blank()) +
    labs(linewidth = "Proportion of Significant Paths",
         col = "Standardized Estimate",
         title = strata) +
    theme(legend.position = "bottom") +
    scico::scale_colour_scico(palette = "hawaii", 
                              midpoint = 0,
                              limits = c(-2, 2))

  ggsave(here::here("outputs", "strata_sem", "aggregated", paste0(strata, ".png")), dpi = 300,
         width = unit(13, "inches"),
         height = unit(11, "inches"),
         bg = "white")
}

map(group_dfs, plot_sem_group)
```

```{r map}
sig_est_df %>% 
  filter(!is.na(sig_est),
         endsWith(lhs, "DHI")) %>%
  group_by(strata, models) %>%
  slice_max(abs(sig_est)) %>%
  select(mod_follows, var, strata, rhs, est)

```

```{r}
ver_model_latent <- '# latent variables
     fs =~ elev_cv + total_biomass + elev_p95
     elev_cv ~ elev_p95
     total_biomass ~ elev_p95

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

hor_model_latent <- '# latent variables
     fs =~ percentage_first_returns_above_2m + basal_area + percentage_first_returns_above_mean
     basal_area ~ percentage_first_returns_above_2m
     percentage_first_returns_above_mean ~ percentage_first_returns_above_2m

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

internal_model_latent <- '# latent variables
     fs =~ elev_cv + elev_stddev

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

basic_model_latent <- '# latent variables
     fs =~ percentage_first_returns_above_2m + elev_p95

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

derived_model_latent <- '# latent variables
     fs =~ loreys_height + basal_area + elev_cv + elev_stddev + gross_stem_volume + total_biomass

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

all_model_latent <- '# latent variables
     fs =~ basal_area + elev_cv + elev_mean + elev_stddev + loreys_height + percentage_first_returns_above_2m + percentage_first_returns_above_mean + total_biomass + gross_stem_volume + elev_p95

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

all_model_latent <- '# latent variables
     fs =~ elev_cv + elev_mean + elev_stddev + loreys_height + percentage_first_returns_above_2m + percentage_first_returns_above_mean +  elev_p95

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

latent_models <- c(all_model_latent, derived_model_latent, basic_model_latent, internal_model_latent, hor_model_latent, ver_model_latent)

```

```{r}

bec_normal <- bec_df %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x)))

splits <- bec_normal %>%
  group_by(bec) %>%
  group_split()

latent_lavaans <- crossing(splits, latent_models) %>%
  mutate(outs = map2(.x = latent_models, .y = splits, .f = sem))

egg <- crossing(splits = list(bec_normal), latent_models) %>%
  mutate(outs = map2(.x = latent_models, .y = splits, .f = sem))
```

```{r}
egg$outs[[6]] %>%
  graph_sem(layout = get_layout(., "layout_nicely"))
```

```{r}
fitmeasures_list <- c("chisq", "df", "pvalue", "cfi", "rmsea", "aic", "srmr")

myfits <- function(lavaan) {
  try(fitmeasures(lavaan, fit.measures = fitmeasures_list), silent = T)
}

tib_out <- latent_lavaans %>%
  mutate(strata = splits %>% map(pull, var = strata) %>% map(unique) %>% unlist(),
         fits = outs %>% map(myfits),
         pars = outs %>% map(parameterestimates) %>% map(tibble)#,
         # interpreted = outs %>% map(effectsize::interpret)
         ) %>%
  relocate(splits, latent_models)

non_converge <- tib_out %>%
  filter(fits %>% map(class) == "try-error")

converge <- tib_out %>%
  # remove any models that don't converge
  filter(!(fits %>% map(class) == "try-error"))

converge_fm <- converge %>%
  select(latent_models, strata, fits, pars) %>%
  mutate(fits = fits %>% map(as.numeric) %>%
           map(tibble) %>%
           map(mutate, measures = fitmeasures_list) %>%
           map(rename, value = `<dbl>`) %>%
           map(pivot_wider, names_from = measures, values_from = value)) %>%
  unnest(fits)
```

```{r}
crossing(df = list(normal_df), latent_models) %>%
  mutate(outs = map2(.x = latent_models, .y = df, .f = sem))
```

```{r}
sem(latent_models[[1]], data = normal_df) %>%
  summary()
```

```{r}
for(s in clean_df %>%
    pull(strata) %>%
    unique) {
  ggally_plot <- clean_df %>%
    filter(strata == s) %>%
    select(-bec,-category) %>%
    select(strata:total_biomass) %>%
    mutate(MinDHI = normal(MinDHI)) %>%
    #select(CumDHI:total_biomass) %>%
    GGally::ggpairs(aes(colour = strata, alpha = 0.2))

  ggsave(
    here::here("outputs", glue::glue("ggally_{s}.png")),
    plot = ggally_plot,
    dpi = 300,
    height = 40,
    width = 40
  )
}

```

```{r}
# efa
efa_data <- bec_normal %>%
  select(basal_area:VarDHI)

efa_fit <- efa(data = efa_data,
               nfactors = 1:2)

summary(efa_fit, nd = 3L, cutoff = 0.2, dot.cutoff = 0.05)
```

```{r}
# cfa
cfa_model <- "vert =~ elev_mean + elev_p95 + elev_stddev
              hor =~ percentage_first_returns_above_2m + percentage_first_returns_above_mean

              CumDHI ~ vert + hor + elev_cv
              VarDHI ~ vert + hor + elev_cv
              MinDHI ~ vert + hor + elev_cv"

cfa_fit <- sem(cfa_model, bec_normal)

summary(cfa_fit)

lavaanPlot(cfa_fit, coefs = T)
```

```{r}
bec_normal <- bec_df %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x)))

all_model_latent <- '# latent variables
     fs =~ elev_cv + elev_mean + *elev_stddev + *loreys_height + percentage_first_returns_above_2m + *percentage_first_returns_above_mean +  *elev_p95

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

all_model_latent <- '# latent variables
     fs =~ elev_cv + elev_p95 + percentage_first_returns_above_2m
# add hierarchical model here

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

subs <- bec_normal %>%
  slice_sample(n = 500)

egg <- sem(all_model_latent, bec_normal) 
egg %>% summary()

egg %>% semPaths(what = "est")
```

```{r}
ver_model_latent <- '# latent variables
     fs =~ elev_cv + total_biomass + elev_p95
     total_biomass ~ elev_p95

   # dhi follows fs
    CumDHI ~ fs
    MinDHI ~ fs
    VarDHI ~ fs
   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

egg <- sem(ver_model_latent, bec_normal)

egg %>% summary()

egg %>% semPaths(what = "est")
```

Heatmaps

```{r}
library(dendextend)
library(cluster)
# clustering to select variables of interest

bec_iter <- bec_normal %>%
  filter(bec != "BG") %>%
  pull(bec) %>%
  unique()

here::here("outputs", "heatmap") %>%
  dir.create()
for (s in bec_iter) {
  spellman <- bec_normal %>%
    filter(strata == s) %>%
    select(basal_area:total_biomass) %>%
    slice_sample(n = 236) %>%
    as.matrix() %>%
    Hmisc::rcorr()

  spellman_p <- spellman %>%
    .$P

  spellman_cor <- spellman %>%
    .$r

  spellman_dist <- as.dist(1 - spellman_cor)

  spellman_tree <- hclust(spellman_dist, method = "complete")

  spellman_dend <- as.dendrogram(spellman_tree)

  plot(color_branches(spellman_dend, k = 4))

  spellman.kmedoids <- pam(spellman_dist, 4)

  kclusters <- spellman.kmedoids$cluster

  kclusters.reordered <- kclusters[order.dendrogram(spellman_dend)]
  dend.colors <-
    unique(get_leaves_branches_attr(color_branches(spellman_dend, k = 4), attr =
                                      "col"))
  plot(branches_attr_by_clusters(spellman_dend, kclusters.reordered , dend.colors))

  color.scheme <-
    rev(RColorBrewer::brewer.pal(10, "RdBu")) # generate the color scheme to use
  color.scheme <- rev(scico::scico(n = 9, palette = "batlow"))

  p_labs <- spellman_p %>%
    cut(
      breaks = c(0, 0.001, 0.01, 0.05, 1),
      labels = c("***", "**", "*", "NS"),
      right = F
    ) %>%
    matrix(nrow = nrow(spellman_p))

  png(
    filename = here::here("outputs", "heatmap", glue::glue("heatmap_{s}.png")),
    height = 12,
    width = 12,
    units = "in",
    res = 300
  )
  gplots::heatmap.2(
    spellman_cor,
    Rowv = ladderize(spellman_dend),
    Colv = ladderize(spellman_dend),
    dendrogram = "both",
    revC = TRUE,
    # rev column order of dendrogram so conforms to natural representation
    trace = "none",
    density.info = "none",
    col = color.scheme,
    key.title = NA,
    key.xlab = "Correlation",
    cellnote = p_labs
    # cexRow = 0.5,
    # cexCol = 0.5
  )

  dev.off()
}
```

```{r}
bec_normal <- bec_df %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x)))

all_model_latent <- '# latent variables
     fs =~ elev_cv + elev_p95 + percentage_first_returns_above_2m

   # dhi follows fs
    CumDHI ~ fs

   # covariances
    CumDHI ~~ MinDHI
    MinDHI ~~ VarDHI
    VarDHI ~~ CumDHI

'

all_models <- c(all_model_latent, models)

bec_normal %>%
  filter(bec != "BG") %>%
  group_by(bec) %>%
  nest() %>%
  crossing(all_models) %>%
  mutate(out = map2(.x = all_models, .y = data, .f = sem))

out <- sem(all_model_latent, data = bec_normal %>%
             filter(bec == "BG"))

summary(out)
```
